---
title: "Frontiers Ne"
author: "Paolo Cozzi"
date: today
toc: true
number-sections: true
format:
  html:
    code-line-numbers: true
    html-math-method: katex
    df-print: paged
    code-fold: true
    code-summary: "Show the code"
    toc-depth: 4
    embed-resources: true  # Embed all resources directly into the HTML file
editor: source
---

```{r setup}
#| label: setup
#| output: false
require(dplyr)
require(tidyr)
require(tibble)
require(memoise)
require(writexl)
require(rcrossref)
require(smarterapi)
```

## Get sheep breeds for Frontiers paper

Search for *Churra* and *Tibetan* sheeps and collect their *code*:

```{r sheep_breeds}
#| label: sheep_breeds
#| code-fold: false
# collect smarter breeds as rows
churra_breed <- smarterapi::get_smarter_breeds(
  query = list(species = "Sheep", name = "Churra")
)
tibetan_breed <- smarterapi::get_smarter_breeds(
  query = list(species = "Sheep", name = "Tibetan")
)

# select code as a single character string
churra_code <- churra_breed %>% 
  dplyr::select(code) %>%
  dplyr::pull()
tibetan_code <- tibetan_breed %>% 
  dplyr::select(code) %>%
  dplyr::pull()
```

## Collect selected samples

Select the desired samples. Use multiple query terms at the same time

```{r sheep_samples}
#| label: sheep_samples
#| code-fold: false
sheep_samples <- smarterapi::get_smarter_samples(
  species = "Sheep",
  query = list(breed_code = churra_code, breed_code = tibetan_code)
)
```

Now collect all datasets, and merge them with samples. Collect a subset of columns
and then group them:

```{r sheep_samples_merged}
#| label: sheep_samples_merged
#| code-fold: false
#| depends: sheep_samples
sheep_datasets <- smarterapi::get_smarter_datasets(
  query = list(species = "Sheep", type = "genotypes")
)
sheep_samples_merged <- dplyr::inner_join(
  sheep_datasets, 
  sheep_samples, 
  by = dplyr::join_by(`_id.$oid` == `dataset_id.$oid`)
)

# collect only the desired columns
columns <- c(
  "_id.$oid",
  "breed_code",
  "chip_name.y",
  "partner",
  "uploader",
  "country.y",
  "type.y",
  "doi"
)

sheep_info <- sheep_samples_merged %>%
  dplyr::select(all_of(columns)) %>%
  dplyr::rename(
    "dataset_id$oid" = "_id.$oid",
    "chip_name" = "chip_name.y",
    "country" = "country.y",
    "type" = "type.y"
  ) %>%
  dplyr::group_by(
    `dataset_id$oid`,
    breed_code,
    chip_name,
    partner,
    uploader,
    country,
    type,
    doi
  ) %>%
  dplyr::summarise(count = n())
sheep_info
```

There are two main references for *Tibetan* and *Churra* datasets. There are also some samples which comes 
as *foreground* data

## Collect DOI information

Let's resolve the doi:

```{r sheep_doi}
#| label: sheep_doi
#| depends: sheep_info
# declare some stuff before resolving dois

# Function to strip the URL prefix if it exists
strip_doi_url <- function(doi) {
  # Use sub() to remove the https://doi.org/ part if it exists
  stripped_doi <- sub("^https?://(dx\\.)?doi\\.org/", "", doi)
  return(stripped_doi)
}

get_first_non_na_author <- function(authors) {
  # Loop through each row of authors (assuming authors is a data frame)
  for (i in seq_len(nrow(authors))) {
    # Access each author as a row (a list or named vector)
    author <- authors[i, ]
    
    # Check if family name exists and is not NA
    if (!is.null(author[["family"]]) && !is.na(author[["family"]])) {
      return(author[["family"]])  # Return the first valid family name
    }
  }
  
  # If no valid family name found, return "Unknown Author"
  return("Unknown Author")
}

# define an helper function to collect DOI information
get_citation_from_doi <- function(doi) {
  # First strip any URL prefix from the DOI
  doi <- strip_doi_url(doi)
  
  # Silently return NA for NA input
  if (is.na(doi)) {
    return(NA)
  }
  
  # If not cached, retrieve metadata and cache the result
  tryCatch({
    # Retrieve metadata from CrossRef
    metadata <- rcrossref::cr_cn(doi, format = "citeproc-json")
    
    # Check if author information exists and use the helper function to get the first valid author
    if (!is.null(metadata$author) && length(metadata$author) > 0) {
      first_author <- get_first_non_na_author(metadata$author)
    } else {
      first_author <- "Unknown Author"
    }
    
    # Extract the year
    if (!is.null(metadata$issued$`date-parts`[[1]][1])) {
      year <- metadata$issued$`date-parts`[[1]][1]
    } else {
      year <- "Unknown Year"
    }
    
    # Create citation in "Author et al. Year" format
    citation <- paste(first_author, "et al.", year)
    
    # Return the citation
    return(citation)
  }, error = function(e) {
    # Handle errors (e.g., invalid DOI)
    return(NA)
  })
}

# Memoizing the function to use cache
get_citation_from_doi_cache <- memoise::memoise(get_citation_from_doi)

# Helper function to get full citation in apa style
get_full_citation_apa <- function(doi) {
  tryCatch({
    # Fetch the full citation in apa style using rcrossref
    full_citation <- rcrossref::cr_cn(doi, format = "text", style = "apa")
    return(full_citation)
  }, error = function(e) {
    # If there's an error (e.g., invalid DOI), return NA
    return(NA)
  })
}

# Memoizing the function to use cache
get_full_citation_apa_cached <- memoise::memoise(get_full_citation_apa)

# now add citation to the sheep info table
citations <- sheep_info %>%
  dplyr::mutate(short_citation = sapply(doi, get_citation_from_doi_cache)) %>%
  dplyr::mutate(short_citation = tidyr::replace_na(short_citation, "unpublished"))
citations
```

Collect the full bibliography:

```{r sheep_bibliography}
#| label: sheep_bibliography
#| warning: false
#| depends: sheep_doi
# Create a data frame with DOIs and short citations
bibliography <-citations %>%
  dplyr::ungroup() %>%
  dplyr::select(short_citation, doi) %>%
  dplyr::distinct() %>%
  dplyr::filter(!is.na(doi))

# Add a column for the full citation using the apa style
bibliography <- bibliography %>%
  dplyr::mutate(full_citation = sapply(doi, get_full_citation_apa_cached)) %>%
  dplyr::select(c("short_citation", "full_citation", "doi")) %>%
  tibble::as_tibble()
bibliography
```

Save stuff into `.xlsx` file:

```{r create_xlsx}
writexl::write_xlsx(
  list(
    citations = citations,
    bibliography = bibliography),
  "sheep_info.xlsx")
```

Show [sheep_info.xlsx](sheep_info.xlsx)
